---
title: "01-Species-Detected-Wilcoxon-Tests"
author: "Pooja Choksi"
date: '2022-11-03'
output: html_document

---
This chunk of code is used to calculate the cumulative number of species detected from manual annotation of randomly selected acoustic data between 5:30 to 9:30 AM over data from 2020 and 2021. Each figure and table in the main manuscript and supplementary information has been referenced in the description of the separate chunks of code.  

#INSTALL ALL LIBRARIES NEEDED FOR ANALAYSIS:

```{r setup, include=DALSE, message = FALSE, warning= FALSE}

#Install all necessary libraries 
library(dplyr)
library(tidyverse)
library(ggplot2)
library(reshape)
library(stringr)
library(interp)
library(ggsignif)
library(ggpubr)
```

#READ IN ALL DATA FOR THE ANALYSIS PRESENTED IN FIGURE 2, TABLES S6 AND TABLE S7.

```{r include= FALSE, message = FALSE, warning=FALSE}

#CSV of manual annotations of 10-second segments (raw data) of randomly selected data for each sampling location (recorder location)
Data_annotation <- read.csv("C:\\Users\\pooja\\Documents\\GitHub\\Restoration-Soundscapes-RestEcol\\Winter2020-2021-data-annotation-merged.csv")

Data_annotation = select(Data_annotation, -X)

#CSV of species habitat preference 
species_habitat <- read.csv("C:\\Users\\pooja\\Documents\\GitHub\\Restoration-Soundscapes-RestEcol\\Bird_Mammal_Species_List_Mandla_20211211.csv")


```

#CALCULATE CUMULATIVE NUMBER OF SPECIES DETECTED (ALSO PER YEAR)
This chunk of code is used to calculate the cumulative number of species detected in each year (winter season) across all sites. 

```{r message=FALSE, warning=FALSE, include=TRUE}

#Sum up all the instances of a species vocalizing by th recorder location
species_per_site <- Data_annotation%>%group_by(Site_Code, Site_Name, Season, Treatment)%>% summarise(across(where(is.numeric), sum))

#Remove all other non-avian species columns to calculate only the cumulative number of bird species per site  
drop.cols <- c("Uniqueid",
               "X ",
               "Insect_sp_1", 
               "Insect_sp_2",
               "Insect_sp_3",
               "Insect_sp_4",
               "Insect_sp_5", 
               "Insect_sp_6",
               "Insect_sp_7",
               "Insect_sp_8","SD_card",
               "Amphibian_Presence",
               "Amphibian_sp_1",
               "Amphibian_sp_2",
               "Amphibian_sp_3", 
               "Human_1_grazing",
               "Human_2_walking_conversing",
               "Human_2._walking_conversing",
               "Human_3_feral_dogs",
               "Human_4_wood_cutting",
               "Human_5_rooster",
               "Human_6_chainsaw",
               "Human_7_vehicular_noise",
               "Human_8_gunshot","Human_9_music_playing",
               "Human_10_alarm","Quality",
               #also remove all the mammals from the data
               "NPLA","RHMA","TSPS","SPDE","BADE", "GOJA", "BLBU", "GAUR", "Rain_Wind","WIBO", "SADE")

species_per_site <-species_per_site%>%dplyr::select(-one_of(drop.cols))

#Remove NAs in columns from manual annotation
species_per_site<-  species_per_site%>%group_by(Site_Code, Site_Name, Treatment, Season)%>%
  dplyr::mutate_all(funs(replace_na(.,0)))

#Using the totals from species_per_Site, we calculate cumulative species detected per year
#first note species presence/ absence in the totals
total_species_detected <- species_per_site %>% mutate_if(is.numeric, ~1 * (. > 0))

#sum all species presences at sampling location for total species detected per year
total_species_detected <- total_species_detected%>%mutate(Site_Species_Detected= rowSums(across(where(is.numeric))))

#add total species richness from the species richness file to species_per_site file 
total_species_detected <- merge(species_per_site, total_species_detected[,c("Site_Species_Detected", "Site_Code", "Season")],by.x= c("Site_Code", "Season"), by.y= c("Site_Code", "Season"))

```

#FIGURE 2(A) CUMULATIVE NUMBER OF SPECIES DETECTED 
This chunk of code is to make the violin plot in Fig 2A of the cumulative number of species deteced in the sites. Wilcoxon test results presented in Tables S6 and S7

```{r message=FALSE, warning=FALSE, include=TRUE}
#FIG 3A: Cumulative number of species detected

#colour blind friendly palette
cbp <- c("#999999", "#E69F00", "#56B4E9")

#Plot Fig 2A 
cum_num_detected = ggplot(total_species_detected, aes(x = Treatment, y = Site_Species_Detected, fill = Treatment)) +
  geom_violin(width = 0.8)+
  geom_boxplot(width=0.1, color="black", alpha=0.2) + 
  scale_fill_manual(values = cbp)+
  theme_classic()+
  scale_y_continuous(name = "Cumulative number of species detected", breaks=seq(0, 80, 10)) + xlab("")+
  scale_x_discrete(labels = c("LLD", "Restored", "Unrestored"))+
  theme(legend.position = "", text = element_text(size=12),
        axis.text.x = element_text(angle=0, hjust=0.5))

```

#CALCULATE CUMULATIVE SPECIES DETECTED AS PER HABITAT PREFERENCES 
This chunk of code is to assign habitat preferences to all the bird species detected in the manually annotated acoustic data. Habitat preference was taken from State of India's Birds 2020. 

```{r include = TRUE, message=FALSE, warning=FALSE}

#Remove the mammals in this list 
species_habitat <- head(species_habitat,-10)

total_species_detected <- subset(total_species_detected, select = -X)

#pivot longer
total_species_detected_long <- total_species_detected%>%
  pivot_longer(!c(Site_Code,Site_Name, Site_Species_Detected, Treatment, Season), names_to = "Species_Code", values_to = "Species_Presence")

#stitch the tables
species_detected_habitat <- merge(total_species_detected_long, species_habitat[,c("Name_Code", "Habitat_specialization",
"Migratory_Status", "Foraging_pref_1", "Foraging_pref_2","Diet_type_1", "Diet_type_2")], by.x = "Species_Code", by.y = "Name_Code")

#Make separate columns for each habitat type 
species_detected_habitat <- species_detected_habitat%>%
  dplyr::mutate(woodland = case_when(Habitat_specialization == "Woodland" &Species_Presence>0 ~1,TRUE~ 0))

species_detected_habitat <- species_detected_habitat%>%
  dplyr::mutate(generalist = case_when(Habitat_specialization == "Generalist" &Species_Presence>0 ~1,TRUE~ 0))

species_detected_habitat <- species_detected_habitat%>%
  dplyr::mutate(grassland = case_when(Habitat_specialization == "Grassland" & Species_Presence>0 ~1,TRUE~ 0))

species_detected_habitat <- species_detected_habitat%>%
  dplyr::mutate(wetland = case_when(Habitat_specialization == "Wetland" &
 Species_Presence>0 ~1,TRUE~ 0))

species_detected_habitat <- species_detected_habitat%>%
  dplyr::mutate(scrub = case_when(Habitat_specialization == "Scrub" & Species_Presence>0 ~1,TRUE~ 0))

species_detected_habitat <- species_detected_habitat%>%
  dplyr::mutate(tropical_forest = case_when(Habitat_specialization == "Tropical Forest" & Species_Presence>0 ~1,TRUE~ 0))

#determine if species is generalist of forest-woodland affiliated 
species_detected_habitat_sum <- species_detected_habitat%>%dplyr::mutate(generalist_or_forest = case_when(generalist == "1" & Species_Presence>0 ~"Generalist", woodland == "1" & Species_Presence>0 ~"Forest-affiliated",
tropical_forest == "1" & Species_Presence>0~ "Forest-affiliated",
TRUE~ "none"))

write.csv(species_detected_habitat, "Winter2020-2021-species-detected-habitat-long.csv")

```

#CALCULATE CUMULATIVE NUMBER OF FOREST AND WOODLAND AFFILIATED SPECIES DETECTED 

```{r include= TRUE, message=FALSE, warning=FALSE}

#Sum up cumulative species detected according to habitat preference
#NOTE: Here one bird (Eurasian Hoopoe - Upupa epops) is a grassland and scrubland bird
#which is not included in the sum for either habitat preference 
#this does not affect the analysis as all scrub/grassland/wetland birds removed from analysis. they account for <5% of the prop of all birds.

species_habitat_sum <- species_detected_habitat%>%
  dplyr::group_by(Site_Code, Site_Name, Site_Species_Detected, Treatment, Season)%>%dplyr::summarise(woodland_birds = sum(woodland), 
            generalist_birds = sum(generalist), 
            grassland_birds = sum(grassland), 
            wetland_birds = sum(wetland), 
            scrub_birds = sum(scrub), 
            tropical_forest_birds = sum(tropical_forest)) 

#Sum up cumulative species detected according to generalist and forest species 
generalist_forest_sum <- species_habitat_sum%>%
  group_by(Site_Code, Site_Name, Site_Species_Detected, Treatment, Season)%>% summarise(forest_birds = sum(woodland_birds, tropical_forest_birds), generalist_birds = sum(generalist_birds))


```

#FIGURE 2(B): CUMULATIVE NUMBER OF GENERALIST SPECIES DETECTED

This chunk of codes takes the generalist and forest birds sum to plot the cumulative number of generalists detected in the different site types. Wilcoxon test results presented in Tables S6 and S7

```{r message=FALSE, warning=FALSE, include=TRUE}

cum_generalist_detected = ggplot(generalist_forest_sum, aes(x = Treatment, y = generalist_birds, fill = Treatment)) +
  geom_violin(width = 0.8)+
  geom_boxplot(width=0.1, color="black", alpha=0.2) + 
  scale_fill_manual(values = cbp)+
  theme_classic()+ylab("")+
  scale_y_continuous(breaks=seq(0, 80, 10)) +
  scale_x_discrete(labels = c("LLD", "Restored", "Unrestored"))+
  theme(legend.position = "", text = element_text(size=12),
        axis.text.x = element_text(angle=0, hjust=0.5))

```

#FIGURE 2(C): CUMULATIVE FOREST AND WOODLAND AFFILIATED BIRDS DETECTED 
This chunk of codes takes the generalist and forest birds sum to plot the cumulative number of forest and woodland affiliated birds detected in the different site types. Wilcoxon test results presented in Tables S6 and S7

```{r message=FALSE, warning=FALSE, include=TRUE}

cum_forest_detected = ggplot(generalist_forest_sum, aes(x = Treatment, y = forest_birds, fill = Treatment)) +
  geom_violin(width = 0.8)+
  geom_boxplot(width=0.1, color="black", alpha=0.2) + 
  scale_fill_manual(values = cbp)+
  theme_classic()+ ylab("")+ xlab("")+
  scale_y_continuous(breaks=seq(0, 30, 10)) +
  scale_x_discrete(labels = c("LLD", "Restored", "Unrestored"))+
  theme(legend.position = "", text = element_text(size=12),
        axis.text.x = element_text(angle=0, hjust=0.5))

```

#COMBINE ALL PLOTS FOR FIGURE 2
This chunk of code puts together the three plots created above for the final figure 2 in the main manuscript. 

```{r include= TRUE, message=FALSE, warning=FALSE}

ggarrange(cum_num_detected, cum_generalist_detected, cum_forest_detected,
          labels = c("A", "B", "C"),
          nrow = 1,
          heights = c(9, 9),
          widths = c(8,8),
          align = "h")

```

#SAVE CSV OF CUMULATIVE NUMBER OF GENERALIST AND FOREST AND WOODLAND AFFILIATED SPECIES AND TOTAL NUMBER OF SPECIES DETECTED

```{r message=FALSE, warning=FALSE, include=FALSE}

write.csv(generalist_forest_sum, "Winter2020-2021-generalist-forest-sum.csv")
```

#WILCOXON TEST RESULTS 
This chunk of code creates the piared datasets used for the Wilcoxon tests which are presented in Tables S6 and Table S7

```{r message=FALSE, warning=FALSE, include=TRUE}

#Create the datasets of different pairs: restored - unrestored, restored- LLD, and unrestored- LLD
to_remove <- generalist_forest_sum %>%filter(Treatment == "No_Lantana")
habitat_specialisation_res_unres <- anti_join(generalist_forest_sum, to_remove)

to_remove <- generalist_forest_sum %>%filter(Treatment == "Unrestored")
habitat_specialisation_res_bench <- anti_join(generalist_forest_sum, to_remove)

to_remove <- generalist_forest_sum%>%filter(Treatment == "Restored")
habitat_specialisation_unres_bench <- anti_join(generalist_forest_sum, to_remove)

#year wise dataset for restored- unrestored
habitat_specialisation_res_unres_2020 <- habitat_specialisation_res_unres%>%filter(Season == "Winter_2020")
habitat_specialisation_res_unres_2021 <- habitat_specialisation_res_unres%>%filter(Season == "Winter_2021")

#year wise dataset for restored- LLD
habitat_specialisation_res_bench_2020 <- habitat_specialisation_res_bench%>%filter(Season == "Winter_2020")
habitat_specialisation_res_bench_2021 <- habitat_specialisation_res_bench%>%filter(Season == "Winter_2021")

#year wise dataset for unrestored- LLD
habitat_specialisation_unres_bench_2020 <- habitat_specialisation_unres_bench%>%filter(Season == "Winter_2020")
habitat_specialisation_unres_bench_2021 <- habitat_specialisation_unres_bench%>%filter(Season == "Winter_2021")

```

#TABLE S6: WILCOXON TEST RESULTS FOR BOTH YEARS
This chunk of code is used ofr the Wilcoxon tests, the results of which are present in Table S6.

```{r include=TRUE, message=FALSE, warning=FALSE}

##Table S6(A): Cumulative number of species detected
##Restored- Unrestored
wilcox_both_years <- wilcox.test(Site_Species_Detected ~ Treatment, data = habitat_specialisation_res_unres, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_both_years$p.value/2)
Zstat
wilcox_both_years

#For the median and quantile values to be reported
res_unres_birds_median <- habitat_specialisation_res_unres%>%
  group_by(Treatment)%>%summarise(median_birds = median(Site_Species_Detected), lower_quantile = quantile(Site_Species_Detected, 0.25),upper_quantile = quantile(Site_Species_Detected, 0.75))
res_unres_birds_median

##Restored- LLD
wilcox_both_years <-wilcox.test(Site_Species_Detected ~ Treatment, data =habitat_specialisation_res_bench, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_both_years$p.value/2)
Zstat
wilcox_both_years 

#For the median and quantile values to be reported
res_bench_birds_median <- habitat_specialisation_res_bench%>%
  group_by(Treatment)%>%summarise(median_birds = median(Site_Species_Detected), lower_quantile = quantile(Site_Species_Detected, 0.25),upper_quantile = quantile(Site_Species_Detected, 0.75))
res_bench_birds_median

##Unrestored -LLD
wilcox_both_years <-wilcox.test(Site_Species_Detected ~ Treatment, data = habitat_specialisation_unres_bench, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_both_years$p.value/2)
Zstat
wilcox_both_years

#For the median and quantile values to be reported
unres_bench_birds_median <- habitat_specialisation_unres_bench%>%
  group_by(Treatment)%>%summarise(median_birds = median(Site_Species_Detected), lower_quantile = quantile(Site_Species_Detected, 0.25),upper_quantile = quantile(Site_Species_Detected, 0.75))
unres_bench_birds_median


##Table S6(B) Cumulative number of forest and woodland affiliated birds detected
##Restored- Unrestored
wilcox_both_years <- wilcox.test(forest_birds ~ Treatment, data = habitat_specialisation_res_unres, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_both_years$p.value/2)

#Calculate the effect size:
abs(Zstat)/sqrt(44)
wilcox_both_years 
Zstat

#For the median and quantile values to be reported
res_unres_forest_birds_median <- habitat_specialisation_res_unres%>%
  group_by(Treatment)%>%summarise(median_forest_birds = median(forest_birds), lower_quantile = quantile(forest_birds, 0.25),
upper_quantile = quantile(forest_birds, 0.75))
res_unres_forest_birds_median

##Restored - LLD 
wilcox_both_years <- wilcox.test(forest_birds ~ Treatment, data = habitat_specialisation_res_bench, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_both_years $p.value/2)
Zstat
wilcox_both_years

#For the median and quantile values to be reported
res_bench_forest_birds_median <- habitat_specialisation_res_bench%>%
  group_by(Treatment)%>%summarise(median_forest_birds = median(forest_birds), lower_quantile = quantile(forest_birds, 0.25),
upper_quantile = quantile(forest_birds, 0.75))
res_bench_forest_birds_median

##Unrestored - LLD
wilcox_both_years <- wilcox.test(forest_birds ~ Treatment, data = habitat_specialisation_unres_bench, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_both_years $p.value/2)
Zstat
wilcox_both_years

#For the median and quantile values to be reported
unres_bench_forest_birds_median <- habitat_specialisation_unres_bench%>%
  group_by(Treatment)%>%summarise(median_forest_birds = median(forest_birds), lower_quantile = quantile(forest_birds, 0.25),
upper_quantile = quantile(forest_birds, 0.75))
unres_bench_forest_birds_median


##Table S6(C) Cumulative number of generalist birds detected
##Restored- Unrestored
wilcox_both_years <- wilcox.test(generalist_birds ~ Treatment, data = habitat_specialisation_res_unres, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_both_years$p.value/2)
Zstat
wilcox_both_years

#For the median and quantile values to be reported
res_unres_generalist_birds_median <- habitat_specialisation_res_unres%>%
  group_by(Treatment)%>%summarise(median_generalist_birds = median(generalist_birds), lower_quantile = quantile(generalist_birds, 0.25),upper_quantile = quantile(generalist_birds, 0.75))
res_unres_generalist_birds_median

##Restored - LLD
wilcox_both_years <- wilcox.test(generalist_birds ~ Treatment, data = habitat_specialisation_res_bench, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_both_years$p.value/2)
Zstat
wilcox_both_years

#For the median and quantile values to be reported
res_bench_generalist_birds_median <- habitat_specialisation_res_bench%>%
  group_by(Treatment)%>%summarise(median_generalist_birds = median(generalist_birds), lower_quantile = quantile(generalist_birds, 0.25),upper_quantile = quantile(generalist_birds, 0.75))
res_bench_generalist_birds_median

##Unrestored- LLD
wilcox_both_years <- wilcox.test(generalist_birds ~ Treatment, data = habitat_specialisation_unres_bench, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_both_years$p.value/2)
Zstat
wilcox_both_years

#For the median and quantile values to be reported
unres_bench_generalist_birds_median <- habitat_specialisation_unres_bench%>%
  group_by(Treatment)%>%summarise(median_generalist_birds = median(generalist_birds), lower_quantile = quantile(generalist_birds, 0.25),upper_quantile = quantile(generalist_birds, 0.75))
unres_bench_generalist_birds_median

```

#TABLE S7 (A): CUMULATIVE NUMBER OF BIRD SPECIES DETECTED IN EACH YEAR  
This chunk of code runs the Wilcoxon tests for the results presented in Table S7(A).

```{r message = FALSE, warning=FALSE, include=TRUE}
##Part 1: Cumulative number of species detected in 2020
##Restored - Unrestored
wilcox_test_2020 <- wilcox.test(Site_Species_Detected ~ Treatment, data = habitat_specialisation_res_unres_2020, 
                                exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2020$p.value/2)
Zstat
wilcox_test_2020

#For the median and quantile values to be reported
res_unres_birds_median_2020 <- habitat_specialisation_res_unres_2020%>%
  group_by(Treatment)%>%summarise(median_birds = median(Site_Species_Detected), lower_quantile = quantile(Site_Species_Detected, 0.25),upper_quantile = quantile(Site_Species_Detected, 0.75))
res_unres_birds_median_2020

##Restored - LLD 
wilcox_test_2020 <- wilcox.test(Site_Species_Detected ~ Treatment, data = habitat_specialisation_res_bench_2020, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2020$p.value/2)
Zstat
wilcox_test_2020

#For the median and quantile values to be reported
res_bench_birds_median_2020 <- habitat_specialisation_res_bench_2020%>%
  group_by(Treatment)%>%summarise(median_birds = median(Site_Species_Detected), lower_quantile = quantile(Site_Species_Detected, 0.25),upper_quantile = quantile(Site_Species_Detected, 0.75))
res_bench_birds_median_2020

##Unrestored - LLD 
wilcox_test_2020 <- wilcox.test(Site_Species_Detected ~ Treatment, data = habitat_specialisation_unres_bench_2020, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2020$p.value/2)
Zstat
wilcox_test_2020

#For the median and quantile values to be reported
unres_bench_birds_median_2020 <- habitat_specialisation_unres_bench_2020%>%
  group_by(Treatment)%>%summarise(median_birds = median(Site_Species_Detected), lower_quantile = quantile(Site_Species_Detected, 0.25),upper_quantile = quantile(Site_Species_Detected, 0.75))
unres_bench_birds_median_2020

##Part 2: Cumulative number of species detected in 2021
##Restored - Unrestored
wilcox_test_2021 <- wilcox.test(Site_Species_Detected ~ Treatment, data = habitat_specialisation_res_unres_2021, 
                                exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2021$p.value/2)
Zstat
wilcox_test_2021

#For the median and quantile values to be reported
res_unres_birds_median_2021 <- habitat_specialisation_res_unres_2021%>%
  group_by(Treatment)%>%summarise(median_birds = median(Site_Species_Detected), lower_quantile = quantile(Site_Species_Detected, 0.25),upper_quantile = quantile(Site_Species_Detected, 0.75))
res_unres_birds_median_2021

##Restored - LLD 
wilcox_test_2021 <- wilcox.test(Site_Species_Detected ~ Treatment, data = habitat_specialisation_res_bench_2021, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2021$p.value/2)
Zstat
wilcox_test_2021

#For the median and quantile values to be reported
res_bench_birds_median_2021 <- habitat_specialisation_res_bench_2021%>%
  group_by(Treatment)%>%summarise(median_birds = median(Site_Species_Detected), lower_quantile = quantile(Site_Species_Detected, 0.25),upper_quantile = quantile(Site_Species_Detected, 0.75))
res_bench_birds_median_2021

##Unrestored - LLD 
wilcox_test_2021 <- wilcox.test(Site_Species_Detected ~ Treatment, data = habitat_specialisation_unres_bench_2021, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2021$p.value/2)
Zstat
wilcox_test_2021

#For the median and quantile values to be reported
unres_bench_birds_median_2021 <- habitat_specialisation_unres_bench_2021%>%
  group_by(Treatment)%>%summarise(median_birds = median(Site_Species_Detected), lower_quantile = quantile(Site_Species_Detected, 0.25),upper_quantile = quantile(Site_Species_Detected, 0.75))
unres_bench_birds_median_2021

```

#TABLE S7 (B): CUMULATIVE NUMBER OF FOREST AND WOODLAND SPECIES DETECTED IN EACH YEAR 
This chunk of code runs the Wilcoxon tests for the results presented in Table S7(B).

```{r include= TRUE, message=FALSE, warning=FALSE}

##Part 1: Forest and woodland affiliated birds detected in 2020
##Restored - Unrestored 
wilcox_test_2020 <- wilcox.test(forest_birds ~ Treatment, data = habitat_specialisation_res_unres_2020,exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2020$p.value/2)
wilcox_test_2020

#For the median and quantile values to be reported
res_unres_forest_birds_median_2020 <- habitat_specialisation_res_unres_2020%>%
  group_by(Treatment)%>%summarise(median_forest_birds = median(forest_birds), lower_quantile = quantile(forest_birds, 0.25),
upper_quantile = quantile(forest_birds, 0.75))
res_unres_forest_birds_median_2020

##Restored - LLD 
wilcox_test_2020 <- wilcox.test(forest_birds ~ Treatment, data = habitat_specialisation_res_bench_2020,exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2020$p.value/2)
wilcox_test_2020

#For the median and quantile values to be reported
res_bench_forest_birds_median_2020 <- habitat_specialisation_res_bench_2020%>%
  group_by(Treatment)%>%summarise(median_forest_birds = median(forest_birds), lower_quantile = quantile(forest_birds, 0.25),
upper_quantile = quantile(forest_birds, 0.75))
res_bench_forest_birds_median_2020

##Unrestored- LLD 
wilcox_test_2020 <- wilcox.test(forest_birds ~ Treatment, data = habitat_specialisation_unres_bench_2020,exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2020$p.value/2)
wilcox_test_2020

#For the median and quantile values to be reported
unres_bench_forest_birds_median_2020 <- habitat_specialisation_unres_bench_2020%>%
  group_by(Treatment)%>%summarise(median_forest_birds = median(forest_birds), lower_quantile = quantile(forest_birds, 0.25),
upper_quantile = quantile(forest_birds, 0.75))
unres_bench_forest_birds_median_2020

##Part 2: Forest and woodland affiliated birds detected in 2021
##Restored - Unrestored 
wilcox_test_2021 <- wilcox.test(forest_birds ~ Treatment, data = habitat_specialisation_res_unres_2021,exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2021$p.value/2)
wilcox_test_2021

#For the median and quantile values to be reported
res_unres_forest_birds_median_2021 <- habitat_specialisation_res_unres_2021%>%
  group_by(Treatment)%>%summarise(median_forest_birds = median(forest_birds), lower_quantile = quantile(forest_birds, 0.25),
upper_quantile = quantile(forest_birds, 0.75))
res_unres_forest_birds_median_2021

##Restored - LLD 
wilcox_test_2021 <- wilcox.test(forest_birds ~ Treatment, data = habitat_specialisation_res_bench_2021,exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2021$p.value/2)
wilcox_test_2021

#For the median and quantile values to be reported
res_bench_forest_birds_median_2021 <- habitat_specialisation_res_bench_2021%>%
  group_by(Treatment)%>%summarise(median_forest_birds = median(forest_birds), lower_quantile = quantile(forest_birds, 0.25),
upper_quantile = quantile(forest_birds, 0.75))
res_bench_forest_birds_median_2021

##Unrestored- LLD 
wilcox_test_2021 <- wilcox.test(forest_birds ~ Treatment, data = habitat_specialisation_unres_bench_2021,exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2021$p.value/2)
wilcox_test_2021

#For the median and quantile values to be reported
unres_bench_forest_birds_median_2021 <- habitat_specialisation_unres_bench_2021%>%
  group_by(Treatment)%>%summarise(median_forest_birds = median(forest_birds), lower_quantile = quantile(forest_birds, 0.25),
upper_quantile = quantile(forest_birds, 0.75))
unres_bench_forest_birds_median_2021
```

#TABLE S7 (C): CUMULATIVE NUMBER OF GENERALIST SPECIES DETECTED IN EACH YEAR 
This chunk of code runs the Wilcoxon tests for the results presented in Table S7(C).

```{r message= FALSE, warning=FALSE, include= TRUE}
##Part 1: Cumulative number of generalist birds detected in 2020 
##Restored - Unrestored
wilcox_test_2020 <- wilcox.test(generalist_birds ~ Treatment, data = habitat_specialisation_res_unres_2020, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2020$p.value/2)
Zstat
wilcox_test_2020

#For the median and quantile values to be reported
res_unres_generalist_median_2020 <- habitat_specialisation_res_unres_2020%>%
  group_by(Treatment)%>%summarise(median_birds = median(generalist_birds), lower_quantile = quantile(generalist_birds, 0.25),upper_quantile = quantile(generalist_birds, 0.75))
res_unres_generalist_median_2020

##Restored - LLD 
wilcox_test_2020 <- wilcox.test(generalist_birds ~ Treatment, data = habitat_specialisation_res_bench_2020, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2020$p.value/2)
Zstat
wilcox_test_2020

#For the median and quantile values to be reported
res_bench_generalist_median_2020 <- habitat_specialisation_res_bench_2020%>%
  group_by(Treatment)%>%summarise(median_birds = median(generalist_birds), lower_quantile = quantile(generalist_birds, 0.25),upper_quantile = quantile(generalist_birds, 0.75))
res_bench_generalist_median_2020

##Unrestored - LLD 
wilcox_test_2020 <- wilcox.test(generalist_birds ~ Treatment, data = habitat_specialisation_unres_bench_2020, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2020$p.value/2)
Zstat
wilcox_test_2020

#For the median and quantile values to be reported
unres_bench_generalist_median_2020 <- habitat_specialisation_unres_bench_2020%>%
  group_by(Treatment)%>%summarise(median_birds = median(generalist_birds), lower_quantile = quantile(generalist_birds, 0.25),upper_quantile = quantile(generalist_birds, 0.75))
unres_bench_generalist_median_2020

##Part 2: Cumulative number of generalist birds detected in 2021 
##Restored - Unrestored
wilcox_test_2021 <- wilcox.test(generalist_birds ~ Treatment, data = habitat_specialisation_res_unres_2021, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2021$p.value/2)
Zstat
wilcox_test_2021

#For the median and quantile values to be reported
res_unres_generalist_median_2021 <- habitat_specialisation_res_unres_2021%>%
  group_by(Treatment)%>%summarise(median_birds = median(generalist_birds), lower_quantile = quantile(generalist_birds, 0.25),upper_quantile = quantile(generalist_birds, 0.75))
res_unres_generalist_median_2021

##Restored - LLD 
wilcox_test_2021 <- wilcox.test(generalist_birds ~ Treatment, data = habitat_specialisation_res_bench_2021, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2021$p.value/2)
Zstat
wilcox_test_2021

#For the median and quantile values to be reported
res_bench_generalist_median_2021 <- habitat_specialisation_res_bench_2021%>%
  group_by(Treatment)%>%summarise(median_birds = median(generalist_birds), lower_quantile = quantile(generalist_birds, 0.25),upper_quantile = quantile(generalist_birds, 0.75))
res_bench_generalist_median_2021

##Unrestored - LLD 
wilcox_test_2021 <- wilcox.test(generalist_birds ~ Treatment, data = habitat_specialisation_unres_bench_2021, exact = FALSE, conf.int = TRUE)

#Calculate the standardised z statistic Z and call it Zstat
Zstat<-qnorm(wilcox_test_2021$p.value/2)
Zstat
wilcox_test_2021

#For the median and quantile values to be reported
unres_bench_generalist_median_2021 <- habitat_specialisation_unres_bench_2021%>%
  group_by(Treatment)%>%summarise(median_birds = median(generalist_birds), lower_quantile = quantile(generalist_birds, 0.25),upper_quantile = quantile(generalist_birds, 0.75))
unres_bench_generalist_median_2021

```





