---
title: "04-ASU-Species-PERMANOVA"
author: "Pooja Choksi and Mayuri Kotian"
date: '2022-11-08'
output: html_document
---

This chunk of code is used to carry out a PERMANOVA for two outcome variables:
(1) Cumulative number of species detected
(2) Acoustic space used in the 2000-8000 Hz range 

The results of PERMANOVA for each of these outcome variables is presented in Table S8 and Table S13. 

#INSTALL ALL LIBRARIES
```{r message=FALSE, include=FALSE, warning=FALSE}
library(vegan)
library(viridis)
library(dplyr)
library(tidyverse)

```

#READ IN ALL DATA
```{r message=FALSE,include=TRUE, warning=FALSE}

#CSV of Acoustic space use 
asu_low = read.csv("Winter2020-2021-ASU-Prop-2-8k.csv")

#CSV of Species presence and absence 
species_presence_absence = read.csv("Winter2020-2021-sp-pres-abs.csv")

#CSV of predictor variables 
Exp_vars <- read.csv("Vegetation-Census-Geographic-Predictors.csv")

```

#PERMANOVA OF BIRD SPECIES PRESENCE ABSENCE
This chunk of code is used to run a PERMANOVA on the bird community based on manually annotated presence and absences in 10-second audio clips. The results of this analysis are presented in Table S8.

```{r message= TRUE, warning=FALSE, include=TRUE}

#create the dataframe with species presence absence data and the explanatory variables 
species_predictors = merge(species_presence_absence, Exp_vars, by.x= c("Site_Code", "Treatment", "Site_Name"), by.y = c("Site_Code", "Treatment", "Site_Name"))
  
permanova_species = adonis(species_predictors[ , 6:177] ~Total_Trees+ Treatment + Season+ Site_Name + Total_Big_Trees+ TotalPop_3km +Forest_3km + Farms_3km + Simpson_Index_All_Trees, data= species_predictors, permutations =999)
permanova_species$aov.tab


```

#PERMANOVA OF ACOUSTIC SAPCE USED 

This chunk of code is used to run a PERMANOVA on the ASU. The results of this analysis are presented in Table S13. 

```{r message=FALSE,include=TRUE, warning=FALSE}
set.seed(123456) 

#Recalculating Percent ASU
asu_low$Percent_ASU <- (asu_low$File_Count/asu_low$Total_files)*100 

#Remove columns not required for PERMANOVA analysis and store dataframe as a new object
prop_8k_asu <- subset(asu_low, select = -c(9,10)) 

# Remove duplicate rows if any
prop_8k_asu <- prop_8k_asu[!duplicated(prop_8k_asu),] 

# Calculate average Percent ASU per Time x Freq bin grouped by Year, Treatment and sampling location per day
prop_8k_asu <-prop_8k_asu %>% dplyr::group_by(Treatment, Site_Code, Year, Time_24, Frequency_128) %>% dplyr::summarise(Percent_ASU = mean(Percent_ASU)) 

# Combining Time and Frequency labels to create a single label representing each Time x Freq bin hereafter
prop_8k_asu$Time_Freq_bin <- paste(prop_8k_asu$Time_24, prop_8k_asu$Frequency_128, sep = "Time-Freq")

#Create a Season (Year) column to merge with the ASU data
Exp_vars_2020 = Exp_vars
Exp_vars_2020$Season <- 2020

Exp_vars_2021 = Exp_vars
Exp_vars_2021$Season <- 2021

Exp_vars = rbind(Exp_vars_2020, Exp_vars_2021)

#Converting the dataframe from long form to wide form for PERMANOVA analysis
prop_8k_asu_wide <- prop_8k_asu %>%
  pivot_wider(!c(Time_24,Frequency_128), names_from = "Time_Freq_bin", values_from = "Percent_ASU") ## Table in wide format for analysis

#Extracting variables with sampling locations and site names to add to dataframe with percent ASU 
SN <- data.frame(Exp_vars$Site_Code, Exp_vars$Site_Name) 

#Renaming columns to match both dataframes
colnames(SN)[1] <- 'Site_Code' 
colnames(SN)[2] <- 'Site_Name'

#Remove duplicate rows, if any
SN <- SN[!duplicated(SN),]

#sampling location codes with site names 
prop_8k_asu_wide <- merge(SN, prop_8k_asu_wide, by.x = "Site_Code", by.y = "Site_Code")

#check for all site names
unique(prop_8k_asu_wide$Site_Name)

#Replace NA values with 0 before input to PERMANOVA
prop_8k_asu_wide[is.na(prop_8k_asu_wide)] = 0

# Computing PERMANOVA results
permanova_2 = adonis(prop_8k_asu_wide[ , 5:(ncol(prop_8k_asu_wide))] ~ Total_Trees + Treatment + Season + Site_Name+ Total_Big_Trees + TotalPop_3km +Forest_3km + Farms_3km + Simpson_Index_All_Trees, data = Exp_vars, permutations = 999)
permanova_2
permanova_2$aov.tab   

```

#END


